Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-10-02 paper.tex
Power output graph

    We replicate the fabrication method and approximate dimensions of \cite{Gu2000} to produce ten streaming cells.
    The internal dimensions of which are \SI{3}{\centi\metre} long, \SI{1}{\centi\metre} wide and heights in the range of \SI{25}{\micro\metre} to \SI{250}{\micro\metre}.
    These dimensions are directly comparable to those of \cite{Gu2000}, where the length and width are the same but the heights span a wider range.
    The channels themselves are made from glass microscope slides (Sail Brand - JIA 7101WT) sectioned into halves.
    Plastic shims (Garlock Colorplast) are epoxied (Selleys Araldite Ultra Clear Resin) between slide halves to create the channels.
    The dimensions at each end of each channel are measured under a microscope to determine the actual channel height and ensure symmetry.

@@

    Measured data obtained from each of the ten streaming cells is presented in Fig.~\ref{fig:cellEfficiency}.
    Here the ratio of streaming voltage to applied pressure is plotted as a function of the internal channel height.
    It is clear from these measurements that narrower channels offer greater electrical conversion efficiency.

    Based on measurements of the \SI{52}{\micro\meter} channel we estimate the amount of electrical power available for harvesting.
    Fig.~\ref{fig:equivalentCircuit} shows the equivalent circuit for a streaming potential cell delivering a streaming current ($I_{S}$).
    The the streaming potential ($V_{S}$) was measured with no load restance ($R_{L}$) due to the use of the E5270 preventing load current being drawn.
    The dimensions of the cell together with the conductivity of the working fluid give the cell an internal resistance ($R_{ch}$) of approximately \SI{30}{\giga\ohm}.
    Using these parameters, the available output power from the cell has been calculated and is presented in Fig.~\ref{fig:cellPowerPressure} with respect to the applied pressuse.

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_cellEfficiency}
        \end{center}
        \caption{Gradient of developed streaming voltage with applied pressure differential versus the channel height of ten fabricated streaming cells.
        The most efficient cell has a channel height of \SI{52}{\micro\meter}.}
        \label{fig:cellEfficiency}
    \end{figure}

@@

@@

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_pressureLoss}
        \end{center}
        \caption{Estimated head loss from a mechanical water meter typically installed in a domestic setting (Kent PSMT 25mm).}
        \label{fig:headloss}
    \end{figure}

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_harvest}
        \end{center}
        \caption{Calculated power dissipation by a typical domestic mechanical water meter (Kent PSMT 25mm) for each of the sample profile events.}
        \label{fig:powerDissipated_meter}
    \end{figure}

@@

    The Kent PSMT 25mm is a water meter typically installed in homes around the Auckland district.~\cite{WatercareNewZealand2014}.
    Figure~\ref{fig:headloss} shows the head loss curve that we have reconstructed based on product specifications for this meter.
    From that curve, we calculate the amount of power dissipated by the Kent PSMT meter during the average washing machine cycle, shower, and toilet flush.
    This data is presented as figure~\ref{fig:powerDissipated_meter}.
    The total energy dissipated within the water meter for each of the water events is:
    \begin{itemize}
    \item \SI{547}{\joule} per load of washing,
    \item \SI{222}{\joule} per shower, and
    \item \SI{24.3}{\joule} per flush of the toilet.
    \end{itemize}

>>>>

    We replicate the approximate dimensions of \cite{Gu2000} to produce ten streaming cells (\SI{245}{\micro\meter}, \SI{178}{\micro\meter}, \SI{161}{\micro\meter}, \SI{125}{\micro\meter}, \SI{106}{\micro\meter}, \SI{75}{\micro\meter}, \SI{71}{\micro\meter}, \SI{56}{\micro\meter}, \SI{52}{\micro\meter}, \SI{26}{\micro\meter}).
    The channels are made from glass microscope slides (Sail Brand - JIA 7101WT) sectioned into halves.
    Plastic shims (Garlock Colorplast) are epoxied (Selleys Araldite Ultra Clear Resin) between slide halves to separate the slides.
    The dimensions at each end of each channel are measured under a microscope to determine the channel height once the channels had set.

@@

    Measured pressure-to-voltage gradients from each of the ten streaming cells is presented in Fig.~\ref{fig:cellEfficiency}.
    We put the spread in data points down to asymmetry in the cells and uncertainty of the internal dimensions.
    When viewed under the microscope the opposite ends of cell heights differing as much as \SI{3}{\micro\meter} in some cases.
    Due to the opacity of the resin used, measurement of the internal cell dimensions was not possible.

    The plot shows a loose relationship between channel height and voltage gradient per Pascal of applied pressure.
    The cell with height \SI{26}{\micro\meter} shows a much lower slope, probably as a result of a reduced flow rate.
    We were unable to measure the flow rate of the \SI{26}{\micro\meter} cell before it burst.

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_cellEfficiency}
        \end{center}
        \caption{Gradient of developed streaming voltage with applied pressure differential versus the channel height of ten fabricated streaming cells.}
        \label{fig:cellEfficiency}
    \end{figure}

@@

    We now measure available power output of the \SI{71}{\micro\meter} streaming cell.
    This cell was chosen as the \SI{52}{\micro\meter} cell burst during operation and the \SI{71}{\micro\meter} cell seemed an appropriate substitute as per Fig.~\ref{fig:cellEfficiency}.
    We measure the streaming voltage, and pressure, while sweeping the amount of electrical current drawn from the device.
    The applied pressure was held at \SI{260}{\kilo\pascal} for the duration of the measurement and the flow rate was previously determined to be \SI{2.05}{\milli\litre\per\second}.

    Fig.~\ref{fig:cellOutput} shows the measured data along with the calculated power.
    The maximum power theorem states that the maximum power

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_cellOutput}
        \end{center}
        \caption{Cell characteristics plotted against drawn current.
        The cell's output }
        \label{fig:cellOutput}
    \end{figure}

    Based on measurements of the \SI{52i}{\micro\meter} channel we estimate the amount of electrical power available for harvesting.
    Fig.~\ref{fig:equivalentCircuit} shows the equivalent circuit for a streaming potential cell delivering a streaming current ($I_{S}$).
    The the streaming potential ($V_{S}$) was measured with no load resistance ($R_{L}$) due to the use of the E5270 preventing load current being drawn.
    The dimensions of the cell together with the conductivity of the working fluid give the cell an internal resistance ($R_{ch}$) of approximately \SI{30}{\giga\ohm}.
    Using these parameters, the available output power from the cell has been calculated and is presented in Fig.~\ref{fig:cellPowerPressure} with respect to the applied pressure.

@@

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_pressureLoss}
        \end{center}
        \caption{Estimated head loss from a mechanical water meter typically installed in a domestic setting.}
        \label{fig:headloss}
    \end{figure}

    \begin{figure}
        \begin{center}
        \includegraphics[width=\linewidth]{graph_harvest}
        \end{center}
        \caption{Calculated power dissipation by a typical domestic mechanical water meter for each of the sample profile events.}
        \label{fig:powerDissipated_meter}
    \end{figure}

@@

    Figure~\ref{fig:headloss} shows the head loss curve from a water meter typically installed at New Zealand homes (Kent PSMT 25mm).~\cite{WatercareNewZealand2014}
    Using this pressure loss curve we calculate the amount of power dissipated by a typical water meter during an average washing machine cycle, shower, and toilet flush.
    Calculated pressure loss due to each of these events is presented as figure~\ref{fig:powerDissipated_meter}.
    The total energy dissipated within the water meter for each of the water events is:
    \begin{itemize}
    \item \SI{547}{\joule} per load of washing,
    \item \SI{222}{\joule} per shower, and
    \item \SI{24.3}{\joule} per flush of the toilet.
    \end{itemize}
